---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Unzip archive and read raw data from file. Overview the dataset.
```{r echo=TRUE}
unzip('activity.zip', overwrite = TRUE)
raw_data <- read.csv('activity.csv', na.strings = 'NA', header = TRUE)
str(raw_data)
```


## What is mean total number of steps taken per day?

Calculate total steps per each days and draw a histogram of the total number of 
steps taken each day.
```{r echo=TRUE}
steps_per_day <- aggregate(steps~date, data=raw_data, sum, na.rm = TRUE)
hist(steps_per_day$steps, main="Steps per day", xlab = "Total steps per day")
```

Calculate mean and median.
```{r echo=TRUE}
spd_mean <- mean(steps_per_day$steps)
spd_median <- median(steps_per_day$steps)
```

Mean of total number of steps taken per date is `r sprintf('%f', spd_mean)`. 
And median is `r spd_median`.


## What is the average daily activity pattern?

Calculate average steps in 5-minute intervals across all days.
```{r echo=TRUE} 
asteps_5min <- aggregate(steps~interval, data=raw_data, mean, na.rm = TRUE)
```

Convert interval names to format _hh:mm_ and add this to data. Draw a plot.
```{r echo=TRUE}
int_names <-sapply(sprintf("%04d", as.integer(asteps_5min$interval)), 
                   function(x){
                       paste(c(substr(x, 1, 2), substr(x, 3, 4)), collapse=':')

                   })

asteps_5min <- cbind(asteps_5min, 'intname' = int_names)

plot(asteps_5min$steps, type='l', main="Average numbers of steps taken", 
     xlab = "5-min intervals", ylab = "Average number of steps", col = "red",
     xaxt = "n")
lab_pos <- seq(from = 0, by = 50, length.out = nrow(asteps_5min)/50)
axis(1, at=lab_pos, labels = asteps_5min$intname[lab_pos + 1])
```

Find interval with the maximum average number of steps.
```{r echo=TRUE}
max_asteps_ind <- 
    (1:nrow(asteps_5min))[asteps_5min$steps == max(asteps_5min$steps)]
max_asteps_start <- asteps_5min$intname[max_asteps_ind]
max_asteps_end <- c(asteps_5min$intname, '00:00')[max_asteps_ind + 1]
```

Interval [`r max_asteps_start`, `r max_asteps_end`) 
contains the maximum number of steps on average across all the days.

## Imputing missing values

Total number of missing values in the dataset is `r sum(is.na(raw_data$steps))`.
Fill __NA__s with mean for this 5-minute interval. Replace __NA__ 
values

```{r echo=TRUE}
filled_data <- raw_data
is_na_ind <- is.na(raw_data$steps)
filled_data$steps[is_na_ind] <- 
    asteps_5min$steps[filled_data$interval[is_na_ind]]
```

Draw a histogram. Calculate mean and median.
```{r echo=TRUE}
filled_spd <- aggregate(steps~date, data = filled_data, sum, na.rm = TRUE)
hist(filled_spd$steps, main="Steps per day", xlab = "Total steps per day")
mean_fspd <- mean(filled_spd$steps)
median_fspd <- median(filled_spd$steps)
```

Mean of filled dataset is `r mean_fspd`.
Median is `r sprintf('%f', median_fspd)`.

Mean and median is differ from the estimates from the first part. Imputing 
missing data enlarge number of observations, decrease mean and median.

## Are there differences in activity patterns between weekdays and weekends?
Set localce to C for English day names. Check first char of day name for S. If
it is TRUE it's weekend. Else weekday. Create new dataset __wddata__ 
with column __wdtype__.

```{r echo=TRUE}
lct <- Sys.getlocale("LC_TIME"); 
Sys.setlocale("LC_TIME", "C")

wdtypes <- ifelse(
    substr(weekdays(as.Date(raw_data$date, '%Y-%m-%d')), 1, 1) == 'S', 
    'weekend', 
    'weekday')
wddata <- cbind(raw_data, wdtype=wdtypes)
Sys.setlocale("LC_TIME", lct)
```

Caclucalte average number of steps in 5 minute intervals. Plot graph with 
lattice package.

```{r echo=TRUE}
wd5min <- aggregate(steps ~ interval + wdtype, data = wddata, sum, na.rm = TRUE)
library(lattice)
xyplot(steps ~ interval | wdtype, data = wd5min, layout=c(1, 2), type = "l")
```

Activity patterns between weekdays and weekends are differ.
